---
title: 'Exercise 1'
author: 'Johannes Breuer & Stefan Jünger'
date: 'Introduction to Survey Data Cleaning Using Tidyverse in R'
---

```{r knitr_init, echo=FALSE, cache=FALSE, include=FALSE}
# custom boxes
knitr::opts_template$set(
  clues = list(box.title = "Clues",
               box.body = list(fill = "#fff9dc", colour = "black"),
               box.header = list(fill = "#ffec8b", colour = "black"),
               box.icon = "fa-search",
               box.collapse = TRUE)
)
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy(position = c("top", "right"))
```

As this is the first set exercises in this workshop, it is relatively easy and short. Its purpose is also to get used to this exercise format.

You can find the solutions for this exercise and all other exercises in the `./solutions` folder in the repo/directory that contains the course materials. You can easily copy code from these solution files by clicking on the small blue clipboard icon in the upper right corner of the solution boxes showing the code.

For working on the exercises, we suggest that you write `R` scripts (one per set of exercises or a combined one for the whole course) and store them in the `./MY_CODE` folder contained in the course materials. This folder already contains a script stub called `my_script.R` which you can use for getting started.

**NB**: For the code from the solutions to work, you have to make sure that your script uses the root directory of the course materials as working directory. You can check the current working directory with `getwd()` and set it with `setwd()` accordingly.
If you work with *RStudio*, you can also simply open the `tidyverse-workshop-esra-2021.Rproj` file. In that case, the correct to the working directory is already set.

That being said, let's delve into the exercises:

```{block, box.title = "1", box.body = list(fill = "white"), box.icon = "fa-star"}
Please install and/or load the `tidyverse` package.
```

```{block, opts.label = "clues"}
We always recommend using the `easypackages` packages, which can be installed with the command `install.packages("easypackages")`. After loading the package with `library(easypackages)` you can load *and* install packages (if necessary) with one command: `easypackages::packages("fancy_package_1", "fancy_package_2", ...)`
```

```{r first.exercise, solution = TRUE}
easypackages::packages(
  "tidyverse"
)
```

```{block, box.title = "2", box.body = list(fill = "white"), box.icon = "fa-star"}
Load the data set named `USAarrests` that is built into `R` and print it.
```

```{block, opts.label = "clues"}
You first need to load the data into your environment using the `data()` command. Then you can either just run the name of the data set as a command or use the `print()` function.
```

```{r second.exercise, solution = TRUE}
data("USArrests")

USArrests
```

```{block, box.title = "3", box.body = list(fill = "white"), box.icon = "fa-star"}
Now, convert the data to a `tibble` and print it. Compare it to the previous output.
```

```{block, opts.label = "clues"}
You need the `as_tibble()` function from the `tibble` package here.
```

```{r third.exercise, solution = TRUE}
USArrests_tibble <-
  tibble::as_tibble(USArrests)

USArrests_tibble
```

```{block, box.title = "4", box.body = list(fill = "white"), box.icon = "fa-star"}
Have a look at the tibble using the function `glimpse()` from the `dplyr` package. Compare it to the previous outputs. 
```

```{block, opts.label = "clues"}
The argument that you need to provide to `glimpse()` is the name of the dataframe/tibble. 
```

```{r fourth.exercise, solution = TRUE}
dplyr::glimpse(USArrests_tibble)

# dplyr::glimpse() provides another method of displaying the data. In such small 
# data sets, in doesn't make a huge difference. However, as the outputs of tibbles are 
# reduced for large data sets, it provides a convenient method of getting a quick 
# glimpse (pun intended) of the data.
```

```{block, box.title = "5", box.body = list(fill = "white"), box.icon = "fa-star"}
Finally, reload the data, convert it to a tibble, and use `dplyr::glimpse()` in one combined `%>%`-workflow.
```

```{block, opts.label = "clues"}
Remember that the data object precedes the function call: `x %>% f(.)`
```

```{r fifth.exercise, solution = TRUE}
data("USArrests")

USArrests %>% 
  tibble::as_tibble() %>% 
  dplyr::glimpse()
```


