---
title: "Introduction to Survey Data Cleaning Using Tidyverse in R"
subtitle: "Wrap-Up"
author: "Johannes Breuer<br />Stefan Jünger"
date: "2021-07-22"
location: "ESRA 2021"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["default", "default-fonts", "workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include = F}
if (!require(easypackages)) install.packages("easypackages")
library(easypackages)

packages("knitr",
         "rmarkdown",
         "tidyverse",
         "gadenbuie/xaringanExtra",
         "gadenbuie/tweetrmd",
         "hadley/emo",
         "sctyner/memer",
         prompt = F)

options(htmltools.dir.version = FALSE,
        htmltools.preserve.raw = FALSE)

opts_chunk$set(echo = TRUE,
               fig.align = "center")

xaringanExtra::use_xaringan_extra(c("tile_view", "clipboard"))
xaringanExtra::use_extra_styles(hover_code_line = TRUE,
                                mute_unhighlighted_code = FALSE)
```

<div class="my-footer">
  <div style="float: left;"><span>`r gsub("<br />", ", ", gsub("<br /><br />|<a.+$", "", metadata$author))`</span></div>
  <div style="float: right;"><span>`r metadata$location`, `r metadata$date`</span></div>
  <div style="text-align: center;"><span>`r gsub(".+<br />", " ", metadata$subtitle)`</span></div>
</div>

---

## Course recap

```{r recap 1, echo = F}
tribble(
  ~"Session", ~"Example content", 
  "Introduction", "What is the tidyverse?, %>% & tibbles ",
  "Data import", "Importing CSV, SPSS, and Stata files",
  "Data wrangling 1", "Renaming, recoding, & selecting variables; filtering cases",
  "Data wrangling 2", "Transforming variables & grouping cases", 
) %>% 
kable(
    format = "html"
  , align = "ll"
  , escape = FALSE
)
```

---

## Where to go from here?

- working with **strings** with [`stringr`](https://stringr.tidyverse.org/index.html)

- working with **factors** with [`forcats`](https://forcats.tidyverse.org/)

- working with **times and dates** with [`lubridate`](https://lubridate.tidyverse.org/)

- working with **labelled data** with [`haven`](https://github.com/tidyverse/haven) and [`sjlabelled`](https://github.com/strengejacke/sjlabelled) (the latter is not associated with the `tidyverse` but still a great package)

- survey data **analysis** with the [`survey` package](https://cran.r-project.org/web/packages/survey/index.html) (also not part of the `tidyverse`)

- **visualization** with [`ggplot2`](https://ggplot2.tidyverse.org/)

- **reporting & publishing** with [`RMarkdown`](https://rmarkdown.rstudio.com)

- **functional programming** with the [`purrr` package](https://purrr.tidyverse.org/)

There are also chapters on all of these topics in [*R for Data Science*](https://r4ds.had.co.nz/index.html) (with the exception of the `sjlabelled` and `survey` packages)

---

## The expanding `tidyverse`

There is a growing list of packages that implement/adapt `tidyverse` principles and workflows for different types of analyses.  

Some examples:

- [`tidymodels`](https://www.tidymodels.org/) for modeling and machine learning using `tidyverse` principles

- [`tidygraph`](https://www.data-imaginist.com/2017/introducing-tidygraph/) for a tidy approach to network analysis

- [`tidytext`](https://cran.r-project.org/web/packages/tidytext/vignettes/tidytext.html) for a [tidy approach to text mining](https://www.tidytextmining.com/)

- [`tidybayes`](https://mjskay.github.io/tidybayes/index.html) for tidy Bayesian analyses

---

##


---

## The 'dark side' of the `tidyverse`


---

## Acknowledgements `r ji("heart")`

All slides were created with the `R` package [`xaringan`](https://github.com/yihui/xaringan) which builds on [`remark.js`](https://remarkjs.com), [`knitr`](http://yihui.name/knitr), and [`RMarkdown`](https://rmarkdown.rstudio.com). The exercises were created with the [`unilur` package](https://github.com/koncina/unilur).  

Please make sure to properly cite all data that you use for your research (archives usually provide suggested citations). Also make sure to cite the free and open-source software (FOSS) that you use, such as `R` and the packages for it. To know how to do that, you can use the function `citation(packagename)` in `R`.

We thank the ESRA 2021 organizers (especially Marieke Haan), and all of you for participating!
