---
title: "Introduction to Survey Data Cleaning Using Tidyverse in R"
subtitle: "Data Import"
author: "Johannes Breuer<br />Stefan Jünger"
date: "2021-07-22"
location: "ESRA 2021"
output:
  xaringan::moon_reader:
    self_contained: true
    lib_dir: libs
    css: ["default", "default-fonts", "../assets/workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include = FALSE}
# load all packages
source("./content/R/load_packages.R")

options(htmltools.dir.version = FALSE,
        htmltools.preserve.raw = FALSE)

opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)

# load course content table
source("./content/R/course_content.R")

options(warn = -1)

xaringanExtra::use_xaringan_extra(c("tile_view", "clipboard"))
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = FALSE  
)
```

<div class="my-footer">
<div style="float: left;"><span>`r gsub("<br />", ", ", gsub("<br /><br />|<a.+$", "", rmarkdown::metadata$author))`</span></div>
<div style="float: right;"><span>`r rmarkdown::metadata$location`, `r rmarkdown::metadata$date`</span></div>
<div style="text-align: center;"><span>`r gsub(".+<br />", " ", rmarkdown::metadata$subtitle)`</span></div>
</div>

---

## Getting data into `R`

.large[
This course is about cleaning survey data.
- thus, making them tidy

There's one important prerequisite:

.center[**We need data!**]
]

```{r, out.width = "50%", echo = FALSE}
include_picture("import_data.png")
```

---

## Getting data into the `tidyverse`

`base R` already provides tools to import data
- e.g., `read.csv()` to import Comma Seperated Value (CSV) files

And the `tidyverse` provides some nice tools to do that too:
- `readr`  (for 'flat' files, such as CSV)
- `haven`  (for statistical software files, such as SPSS, Stata, and SAS; can also handle labelled data)
- `readxl` (for Excel spreadsheets)

```{r, out.width = "50%", echo = FALSE}
include_picture("hex_all_in_one.png")
```

A key difference between these packages and others is that `tidyverse` import functions already prepare the data for being 'tidied': they are a `tibbles`

---

## A readr example: `read_csv()`
```{r readr_example, echo = TRUE, message = TRUE}
gpc_csv <- 
  readr::read_csv("./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv")
```

Note the column specifications. `readr` 'guesses' them based on the first 1000 observations (we will come back to this later).



---

## A readr example: `read_csv()`
```{r readr-example-print, eval = FALSE}
gpc_csv
```

.tiny[
```{r readr-example-output, echo = FALSE}
gpc_csv
```
]

The *GESIS Panel* website provides [detailed documentation](https://www.gesis.org/gesis-panel/documentation), including a [cheatsheet](https://www.gesis.org/fileadmin/upload/forschung/programme_projekte/Drittmittelprojekte/GESIS_Panel/gesis_panel_cheatsheet.pdf).

---

## A readxl example: `read_excel()`
```{r readxl_example, echo = TRUE}
gpc_xlsx <- 
  readxl::read_xlsx(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.xlsx"
  )
```

No output `r ji("frowning_face")`

---

## A readxl example: `read_excel()`
```{r readxl_example_output, echo = TRUE, eval = FALSE}
gpc_xlsx
```

.tiny[
```{r readr-readxl_example_output-print, echo = FALSE}
gpc_xlsx
```
]

---

## A haven example: `read_stata()` 
```{r read_stata_example, echo = TRUE}
gpc_dta <- 
  haven::read_stata(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.dta"
  )
```

Note: The [`gesisdata` package](https://cran.r-project.org/web/packages/gesisdata/index.html) allows direct access to the *GESIS* data `R`, given that you have [proper credentials](https://login.gesis.org/).

---

## A haven example: `read_stata()` 
```{r read_stata_example_output, echo = TRUE, eval = FALSE}
gpc_dta
```

.tiny[
```{r readr-read_stata_example_output-print, echo = FALSE}
gpc_dta
```
]

---

class: middle

.center[
```{r same-picture-code, eval = FALSE}
gpc_csv  ...  pc_xlsx  ...  gpc_dta
```
]

.pull-left[
.tinylicious[
```{r same-picture-code-print, echo = FALSE}
gpc_csv
```
]
]

.pull-right[
```{r same-picture, echo = FALSE}
include_picture("same_picture.jpg")
```
]


---

## `read_stata()`'s siblings

Indeed, there's also the function `read_spss()` and `read_sas()` to import SPSS and SAS files.

`read_spss()` also provides capabilities to handle SPSS-defined missing values by setting the option `user_na = TRUE` (default is `FALSE`).

The [`sjlabelled` package](https://cran.r-project.org/web/packages/sjlabelled/index.html) can also be used to choose a more elaborated approach for missing values: https://cran.r-project.org/web/packages/sjlabelled/vignettes/intro_sjlabelled.html

---

## There's more

These were just some very first examples of applying functions from each package. They comprise even more functions for different data types.

.pull-left[
`readr`
- `read_csv()`
- `read_tsv()`
- `read_delim()`
- `read_fwf()`
- `read_table()`
- `read_log()`
]

.pull-right[
`haven`
- `read_sas()`
- `read_spss()`
- `read_stata()`
]

Not to mention all the helper functions and options. For example, we can define the cells to read from an Excel file by specifying the option `range = "C1:E4"` in `readxl::read_excel()`

---

## Exporting data

Sometimes our data have to leave the `tidyverse` (or even `R`), for example, if we....

- share data with colleagues who do not use `R`
- want to continue where we left off
- particularly, if data wrangling took a long time

For such purposes we also need a way to export our data.

All of the packages we have discussed in this session also have designated functions for that.

```{r, out.width = "50%", echo = FALSE}
include_picture("export_data.png")
```

---

## Examples: CSV and Stata files

```{r export_csv, echo = TRUE}
readr::write_csv(gpc_dta, "gpc_own.csv")
```

```{r export_stata, echo = TRUE}
haven::write_dta(gpc_csv, "gpc_own.dta")
```

Proof that they have been exported:
```{r list_files, echo = TRUE}
list.files() 
```

```{r echo = FALSE, include = FALSE}
unlink("gpc_own.csv")
unlink("gpc_own.dta")
```

---

# Data types

.pull-left[
characters
- indicated by `<chr>`
- specified by `col_character()`
]

.pull-right[
integers
- indicated by `<int>`
- specified by `col_integer()`
]

.pull-left[
doubles
- indicated by `<dbl>`
- specified by `col_double()`
]

.pull-right[
factors
- indicated by `<fct>`
- specified by `col_factor()`
]

.pull-left[
logical
- indicated by `<lgl>`
- specified by `col_logical()`
]

.pull-right[**There's more, but we'll leave it at that for now.**]

---

## Changing variable types

As mentioned before, `readr::read_csv()` 'guesses' the variable types by scanning the first 1000 observations. NB: **This can go wrong!**

Luckily, we can change the variable type...
- before/while loading the data
- and after loading the data

---

## While loading the data in `readr::read_csv()`

```{r readr_example_col_change, echo = TRUE}
gpc <-
  readr::read_csv(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv",
    col_types = cols(
      age_cat = col_factor()
    )
  )
```

---


```{r readr_example_col_change_2, eval = TRUE}
gpc 
```

.tiny[
```{r readr_example_col_change_display, echo = FALSE}
gpc
```
]

---

# After loading the data

```{r readr_example_col_changeD_after, echo = TRUE}
gpc <- 
  read_csv("./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv")

gpc <-
  gpc %>% 
  readr::type_convert(
    titanic,
    col_types = cols(
      age_cat = col_factor()
    )
  )
```

---

## Additional packages

The great benefit of `tidyverse` import functions is the import of the data as tibbles: the data are potentially tidier.

Several other non-tidyverse packages provide similar benefits as they make use of this universal data format:
- [`sf`](https://github.com/r-spatial/sf) for geospatial data

- [`sjlabelled`](https://cran.r-project.org/web/packages/sjlabelled/index.html) to work with labelled data, e.g., from SPSS or Stata

---

## Other packages for data import

- `base` R

- the [`foreign` package](https://cran.r-project.org/web/packages/foreign/index.html) for SPSS and Stata files

- [`data.table`](https://cran.r-project.org/web/packages/data.table/index.html) or [`fst`](https://www.fstpackage.org/) for large datasets

- [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/index.html) for json files

- [`datapasta`](https://github.com/MilesMcBain/datapasta) for copying and pasting data into tribbles (e.g., from websites, Excel or Word files)

---

class: center, middle

# [Exercise](https://jobreu.github.io/tidyverse-workshop-esra-2021/exercises/Exercise_2.html) time `r ji("weight_lifting_woman")``r ji("muscle")``r ji("running_man")``r ji("biking_man")`

## [Solutions](https://jobreu.github.io/tidyverse-workshop-esra-2021/solutions/Exercise_2.html)