---
title: "Introduction to Survey Data Cleaning Using Tidyverse in R"
subtitle: "Data Import"
author: "Johannes Breuer<br />Stefan Jünger"
date: "2021-07-22"
location: "ESRA 2021"
output:
  xaringan::moon_reader:
    self_contained: true
    lib_dir: libs
    css: ["default", "default-fonts", "../assets/workshop.css"]
    nature:
      highlightStyle: "github"
      highlightLines: true
      countIncrementalSlides: false
---
layout: true

```{r setup, include = FALSE}
# load all packages
source("./content/R/load_packages.R")

options(htmltools.dir.version = FALSE,
        htmltools.preserve.raw = FALSE)

opts_chunk$set(echo = TRUE, fig.align = "center", message = FALSE)

# load course content table
source("./content/R/course_content.R")

options(warn = -1)

xaringanExtra::use_xaringan_extra(c("tile_view", "clipboard"))
xaringanExtra::use_extra_styles(
  hover_code_line = TRUE,         
  mute_unhighlighted_code = FALSE  
)
```

<div class="my-footer">
<div style="float: left;"><span>`r gsub("<br />", ", ", gsub("<br /><br />|<a.+$", "", rmarkdown::metadata$author))`</span></div>
<div style="float: right;"><span>`r rmarkdown::metadata$location`, `r rmarkdown::metadata$date`</span></div>
<div style="text-align: center;"><span>`r gsub(".+<br />", " ", rmarkdown::metadata$subtitle)`</span></div>
</div>

---

## Getting data into `R`

.large[
This course is about cleaning survey data.

There's one important prerequisite for that:

.center[**We need data!**]
]

```{r, out.width = "50%", echo = FALSE}
include_picture("import_data.png")
```

---

## Data for this workshop

For this workshop, we will use a synthetic data set based on the data from the [*GESIS Panel Special Survey on the Coronavirus SARS-CoV-2 Outbreak in Germany*](https://search.gesis.org/research_data/ZA5667). This synthetic data set was created by [Bernd Weiß](https://berndweiss.net/) using the [`synthpop` package](https://www.synthpop.org.uk/).

Original data set:

GESIS Panel Team (2020). *GESIS Panel Special Survey on the Coronavirus SARS-CoV-2 Outbreak in Germany*. GESIS Data Archive, Cologne. ZA5667 Data file Version 1.1.0, [https://doi.org/10.4232/1.13520](https://doi.org/10.4232/1.13520)

*Note*: The [`gesisdata` package](https://cran.r-project.org/web/packages/gesisdata/index.html) allows direct access to *GESIS* data in `R`, given that you have [proper credentials](https://login.gesis.org/).

---

## Getting data into the `tidyverse`

`base R` already provides tools for importing data
- e.g., `read.csv()` for importing Comma Separated Value (CSV) files

The `tidyverse` extends these options through a set of packages:
- `readr`  (for 'flat' files, such as `CSV`)
- `haven`  (for statistical software files, such as *SPSS*, *Stata*, and *SAS*; can also handle labelled data)
- `readxl` (for *Excel* spreadsheets)

```{r, out.width = "50%", echo = FALSE}
include_picture("hex_all_in_one.png")
```

A key difference between these packages and others is that `tidyverse` import functions already prepare the data for being 'tidied' by importing them as `tibbles`.

---

## A readr example: `read_csv()`
```{r readr_example, echo = TRUE, message = TRUE}
gpc_csv <- 
  readr::read_csv("./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv")
```

Note the column specifications. `readr` 'guesses' them based on the first 1000 observations (we will come back to this later).

---

## A readr example: `read_csv()`
```{r readr-example-print, eval = FALSE}
gpc_csv
```

.tiny[
```{r readr-example-output, echo = FALSE}
gpc_csv
```
]

.small[
*Note*: To know what the variables represent, we can consult the [codebook](https://dbk.gesis.org/dbksearch/download.asp?id=67378) for the original data. In general, the *GESIS Panel* website provides very [detailed documentation](https://www.gesis.org/gesis-panel/documentation) for the studies. The [*GESIS Panel* cheatsheet](https://www.gesis.org/fileadmin/upload/forschung/programme_projekte/Drittmittelprojekte/GESIS_Panel/gesis_panel_cheatsheet.pdf) is particularly helpful for understanding the data structure and variable naming scheme.
]

---

## A readxl example: `read_excel()`
```{r readxl_example, echo = TRUE}
gpc_xlsx <- 
  readxl::read_xlsx(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.xlsx"
  )
```

No output `r ji("frowning_face")`

---

## A readxl example: `read_excel()`
```{r readxl_example_output, echo = TRUE, eval = FALSE}
gpc_xlsx
```

.tiny[
```{r readr-readxl_example_output-print, echo = FALSE}
gpc_xlsx
```
]

---

## A haven example: `read_stata()` 
```{r read_stata_example, echo = TRUE}
gpc_dta <- 
  haven::read_stata(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.dta"
  )
```

---

## A haven example: `read_stata()` 
```{r read_stata_example_output, echo = TRUE, eval = FALSE}
gpc_dta
```

.tiny[
```{r readr-read_stata_example_output-print, echo = FALSE}
gpc_dta
```
]

---

class: middle

.center[
```{r same-picture-code, eval = FALSE}
gpc_csv  ...  gpc_xlsx  ...  gpc_dta
```
]

.pull-left[
.tinylicious[
```{r same-picture-code-print, echo = FALSE}
gpc_csv
```
]
]

.pull-right[
```{r same-picture, echo = FALSE}
include_picture("same_picture.jpg")
```
]


---

## `read_stata()`'s siblings

`haven` also offers the functions `read_spss()` and `read_sas()` for importing *SPSS* and *SAS* files.

`read_spss()` also provides capabilities for handling *SPSS*-defined missing values by setting the option `user_na = TRUE` (default is `FALSE`).

We will talk a bit more about missing values in the next session on data wrangling.

---

## But there's more...

These were just some examples of the many functions available for importing data in the `tidyverse` packages. There are many more for various different data types...

.pull-left[
`readr`
- `read_csv()`
- `read_tsv()`
- `read_delim()`
- `read_fwf()`
- `read_table()`
- `read_log()`
]

.pull-right[
`haven`
- `read_sas()`
- `read_spss()`
- `read_stata()`
]

... not to mention all the helper functions and options. For example, we can define the cells to read from an *Excel* file by specifying the option `range = "C1:E4"` in `readxl::read_excel()`. If you want to know more, you can consult the help files for the individual functions.

---

## Exporting data

Sometimes our data have to leave the `tidyverse` (or even `R`), for example, if we....

- share data with colleagues (incl. those who do not use `R`)
- want to continue where we left off
  - particularly, if data wrangling took a long time

All of the packages we have discussed in this session also have designated functions for exporting data.

```{r, out.width = "50%", echo = FALSE}
include_picture("export_data.png")
```

---

## Examples: CSV and Stata files

```{r export_csv, echo = TRUE}
readr::write_csv(gpc_dta, "gpc_own.csv")
```

```{r export_stata, echo = TRUE}
haven::write_dta(gpc_csv, "gpc_own.dta")
```

Proof that they have been exported:
```{r list_files, echo = TRUE}
list.files() 
```

```{r echo = FALSE, include = FALSE}
unlink("gpc_own.csv")
unlink("gpc_own.dta")
```

---

## Data types

As you may know, there are many different data/variable types in `R`, such as...

.pull-left[
characters
- indicated by `<chr>`
- specified by `col_character()`
]

.pull-right[
integers
- indicated by `<int>`
- specified by `col_integer()`
]

.pull-left[
doubles
- indicated by `<dbl>`
- specified by `col_double()`
]

.pull-right[
factors
- indicated by `<fct>`
- specified by `col_factor()`
]

.center[
logical
- indicated by `<lgl>`
- specified by `col_logical()`
]

.center[**There's more, but we'll leave it at that for now.**]

---

## Changing variable types

As mentioned before, `readr::read_csv()` (and other data import functions from the `tidyverse`) 'guesses' the variable types by scanning the first 1000 observations. Of course,this can go wrong.

Luckily, we can change the variable type...
- before/while loading the data
- and after loading the data

---

## While loading the data in `readr::read_csv()`

```{r readr_example_col_change, echo = TRUE}
gpc <-
  readr::read_csv(
    "./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv",
    col_types = cols(
      age_cat = col_factor()
    )
  )
```

---


```{r readr_example_col_change_2, eval = TRUE}
gpc 
```

.tiny[
```{r readr_example_col_change_display, echo = FALSE}
gpc
```
]

---

## After loading the data

```{r readr_example_col_changeD_after, echo = TRUE}
gpc <- 
  read_csv("./data/ZA5667_v1-0-0_Stata14_synthetic-data.csv")

gpc <-
  gpc %>% 
  readr::type_convert(
    gpc,
    col_types = cols(
      age_cat = col_factor()
    )
  )
```

---

## Additional packages

The great benefit of `tidyverse` import functions is the import of the data as tibbles: the data are potentially tidier.

Several other non-tidyverse packages provide similar benefits as they make use of this universal data format:

- [`sjlabelled`](https://cran.r-project.org/web/packages/sjlabelled/index.html) for labelled data, e.g., from *SPSS* or *Stata*

- [`sf`](https://github.com/r-spatial/sf) for geospatial data

---

## Other packages for data import

- `base R`

- the [`foreign` package](https://cran.r-project.org/web/packages/foreign/index.html) for *SPSS* and *Stata* files

- [`data.table`](https://cran.r-project.org/web/packages/data.table/index.html) or [`fst`](https://www.fstpackage.org/) for large data sets

- [`jsonlite`](https://cran.r-project.org/web/packages/jsonlite/index.html) for `json` files

- [`datapasta`](https://github.com/MilesMcBain/datapasta) for copying and pasting data into so-called [tribbles](https://tibble.tidyverse.org/reference/tribble.html), which are special versions of tibbles (e.g., from websites, *Excel* or *Word* files)
---

class: center, middle

# [Exercise](https://jobreu.github.io/tidyverse-workshop-esra-2021/exercises/Exercise_2.html) time `r ji("weight_lifting_woman")``r ji("muscle")``r ji("running_man")``r ji("biking_man")`

## [Solutions](https://jobreu.github.io/tidyverse-workshop-esra-2021/solutions/Exercise_2.html)